The R Quintuplet (R$^5$)

### 2017-04-21 main.tex
Code updated

% Code Listing
\usepackage[framemethod=tikz]{mdframed}
\usepackage{listings}
\surroundwithmdframed[
%  hidealllines=true,
  linewidth=0.25pt,
  linecolor=black!50,
  backgroundcolor=black!4,
  innerleftmargin=8pt,
  innertopmargin=3pt,
  innerbottommargin=3pt]{lstlisting}
\lstset{ %
  basicstyle=\small\tt\linespread{0.75},
  language=Python,
  commentstyle=\color{gray},
  columns=fullflexible,
}

@@

Scientific reproducibility and replicability issues have received increased attention in the past few years and computer science is no exception.
One could think that computer science would mostly be shielded from those issues. 
But precisely because it is easy to believe that if a program runs once (and gives expected results) it will do so forever, crucial steps to transform working code into meaningful scientific contributions are rarely undertaken; computer science is victim of reproducibility and replicability issues precisely because it seems impervious to them.

In fact, a program can fail in many different ways at many different levels for many different reasons.
Borrowing the terms coined by Ann Gobble, for your program to survive for some noticeable time, you have to make sure it is re-runnable, repeatable, reproducible, replicate and re-usable.\\

@@

\begin{lstlisting}
# A random walker (R0)
import numpy as np

# randomly choose ten instances of either 1 or -1. 
steps = np.random.choice([-1,+1], 10)
y = steps.cumsum() # compute the cumulative sum (the sum of all preceeding elements)
print y
\end{lstlisting}

Executed, this program would display 
\begin{lstlisting}
[-1  0 -1 -2 -1 -2 -3 -2 -1 -2] # with `steps` being [-1 1 -1 -1  1 -1 -1  1  1 -1]
\end{lstlisting}

@@

Did you ever wonder if you would be able to run a program you wrote some time
ago? Let's consider a program you wrote yesterday. Most probably you can still
run it today, but would this still be true in a week, in a month, in a year or
in a decade? Probably you would legitimately answer positively for weeks
or even months if you're optimistic, but you may have doubts when considering
longer time periods such as years or even decades. Part of the problem is that
computer science is evolving at a very fast pace and you cannot know how the
system, the software and the libraries your program depends on will evolve.
If you want your program to have a tiny chance to be runnable ten years from
now, you better make it clear what are the dependencies in terms of systems,
software and libraries.\\
% TODO: cite Collberg 2014 study
\cite{Collberg:2016}

\begin{lstlisting}
# A random walker (R1)
# Tested with Python 3.6.0 / Numpy 1.12.0
import numpy as np

# Compute random walk for 10 steps
steps = 2*(np.random.uniform(-1, +1, 10) > 0) - 1
y = steps.cumsum()
print(y)
\end{lstlisting}

In our case, the R$^0$ version of our tiny walker uses the print {\em instruction} that is available in Python 2.x but has been deprecated in Python 3 in favor or a  print {\em function}.
This means this program cannot be run with Python 3 and needs to be fixed.
Since Python 2 will soon be discontinued, it may be best to target directly Python 3 and to use the print function (and as a matter of fact, it will work with Python 2).
Another potential problem is the numpy `choice` function that has been introduced only in numpy v1.7.
If for some reason, the installed version of numpy is older, the program will also fail.
One quick fix is to use a simpler way to generate the steps (even if it is less efficient).
Last and most important, we need to specify all the dependencies of our program.

@@

\begin{lstlisting}
# A random walker (R2)
# Tested with Python 3.6 / Numpy 1.12.0
import numpy as np

# Random generator initialization
seed = 1
np.random.seed(seed)

# Compute random walk for 10 steps
steps = np.random.choice([-1,+1], 10)
y = steps.cumsum()
print(y)
\end{lstlisting}

\clearpage
\section*{Reproducible (R$^{\mathbf 3}$)}

@@

\begin{lstlisting}
# A random walker (R3)
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
# Release under the BSD 2 clauses license
# Tested with Python 3.6 / Numpy 1.12.0 using macOS 10.12.4 on a 64 bits architecture
import numpy as np

@@

    steps = 2*(np.random.uniform(-1,+1,n) > 0) - 1
    return steps.cumsum()

if __name__ == '__main__'
    # Unit test
    np.random.seed(1)
    assert walk(10) == [-1, 0, -1, -2, -3, -4, -5, -6, -7, -6]

    # Random generator initialization
    seed = 1
    np.random.seed(seed)

    # Compute random walk for n steps
    y = walk(n=10)

    # Save and display results
    results = {"seed": seed, "y": y}
    print(results["y"])
\end{lstlisting}

@@

\begin{lstlisting}
# A random walker program (R4)
# A replication of the random walker using pure Python
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
# Release under the BSD 2 clauses license
# Tested with Python 3.6 using macOS 10.12.4 on a 64 bits architecture
import random
from itertools import accumulate

def state(seed):
    """ Compute the Mersenne Twister keys from given seed """
    mt = [seed] * (624+1)
    mt[-1] = 624
    for i in range(1, 624):
        mt[i] = int(0xFFFFFFFF & (1812433253 * (mt[i-1]^mt[i-1]>>30)+i))
    return (3, tuple(mt), None)

def walk(n=10):
    """ Random walk for n steps """

if __name__ == '__main__'
    # Unit test
    random.setstate(state(seed=1))
    assert walk(n=10) == [-1, 0, -1, -2, -3, -4, -5, -6, -7, -6]

    # random generator initialization
    seed = 1
    random.setstate(state(seed))

    # Compute random walk for n steps
    y = walk(n=10)

    # Save and display results
    results = {"seed": seed, "y": y}
    print(results["y"])
\end{lstlisting}

@@

>>>>

% Code Listing
\usepackage{listings}
\usepackage[framemethod=tikz]{mdframed}
\makeatletter
\def\mdf@@codeheading{Code Listings}
\define@key{mdf}{title}{\def\mdf@@codeheading{#1}}
\mdfdefinestyle{lstlisting}{%
  backgroundcolor=black!2.5,
  innertopmargin=2pt,
  middlelinewidth=0.75pt,
  outerlinewidth=9pt,
  outerlinecolor=white,
  innerleftmargin=10pt,
  innerrightmargin=10pt,
  leftmargin=0pt,
  rightmargin=0pt,
  rightline=false,
  leftline=false,
  bottomline=false,
  skipabove=\topskip,
  skipbelow=\topskip,
  roundcorner=0pt,
  singleextra={
    \node[text=black, % fill=white, draw,
          anchor=south west, yshift=-0.25pt, xshift=5pt,
          font=\sc\footnotesize] at (O|-P) {\csname mdf@@codeheading\endcsname};},
  firstextra={
    \node[text=black, % fill=white, draw,
          anchor=south west, yshift=-0.5pt, xshift=5pt,
          font=\sc\footnotesize] at (O|-P) {\csname mdf@@codeheading\endcsname};}
}

\lstset{ %
  basicstyle=\small\tt\linespread{0.75},
  language=Python,
  commentstyle=\color{gray},
  columns=fullflexible,
}

\lstnewenvironment{code}[2][]{%
  \lstset{#1}%
  \mdframed[style=lstlisting,title={#2}]%
}{\endmdframed}

%\usepackage{listings}
%\surroundwithmdframed[
%%  hidealllines=true,
%  linewidth=0.25pt,
%  linecolor=black!50,
%  backgroundcolor=black!4,
%  innerleftmargin=8pt,
%  innertopmargin=3pt,
%  innerbottommargin=3pt]{lstlisting}
%\lstset{ %
%  basicstyle=\small\tt\linespread{0.75},
%  language=Python,
%  commentstyle=\color{gray},
%  columns=fullflexible,
%}

@@

Scientific reproducibility and replicability issues have received increased attention in the past few years and computer science is no exception.
One could think that computer science would mostly be shielded from those issues. 
But precisely because it is easy to believe that if a program runs once (and gives expected results) it will do so forever, crucial steps to transform working code into meaningful scientific contributions are rarely undertaken; computer science is plagued by replication problems precisely because it seems impervious to them.\\

In fact, a program can fail in many different ways at many different levels for many different reasons.
Borrowing the terms coined by Carole Anne Gobble, for your program to survive for some noticeable time, you have to make sure it is re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$), replicable (R$^4$) and re-usable (R$^5$).\\

@@

\begin{code}{Random walk (R$^0$)}
# Random walk (R0: worst you can do)
import random

x = 0
for i in xrange(10):
    step = random.choice([-1,+1])
    x = x + step
    print x,
\end{code}
% That would be too easy if the program was commented...
% # randomly choose ten instances of either 1 or -1. 
% # compute the cumulative sum (the sum of all preceeding elements)

Executed, this program would display 
\begin{code}{Output}
[-1  0 -1 -2 -1 -2 -3 -2 -1 -2] # with `steps` being [-1 +1 -1 -1 +1 -1 -1 +1 +1 -1]
\end{code}

@@

Have you ever try to re-run a program you wrote some years ago? Frustratingly, it is often surprisingly hard. 
Part of the problem is that computer science is evolving at a fast pace and you cannot know how the system, the software and the libraries your program depends on will evolve.
Since you wrote the code, you may have changed your computer or your operating system. The compiler, interpreter or set of libraries installed may have changed or been upgraded. 
You may find yourself battling with arcane issues of library compatibility---thoroughly orthogonal to your immediate research goals---to execute again a code \emph{that worked perfectly before}. 
% "thoroughly orthogonal to your immediate research goals": possibly problematic

To be clear, writing code that will run flawlessly months from now---without maintenance---is challenging. When considering years or decades, the best efforts can be stymied by the smallest changes, outside of the programmer's control. And modernizing a unmaintained ten-year-old code can reveal itself to be so complex and expensive that recreating the old execution environment is easier. For this however, the dependencies in terms of systems, software and libraries must be made clear enough.

A \emph{re-runnable} code is one that describes an execution environment in which it is executable and produce the expected results with enough details to be recreated.

% TODO: cite Collberg 2014 study
\cite{Collberg:2016}

\begin{code}{Random walk (R$^1$)}
# Random walk (R2: re-runnable)
# Tested with Python 3
import random
from itertools import accumulate

steps = random.choices([-1,+1], k=10)
x = list(accumulate(steps))
print(x) 
\end{code}

In our case, the R$^0$ version of our tiny walker seems to imply that any version of Python would be fine. This not the case for two reasons.

First, the R$^0$ version uses the print {\em instruction} that is available in Python 2 (this widly used, and installed by default in many operating systems) but has been deprecated in Python 3 (the current version, first released in 2008, almost a decade ago) in favor or a  print {\em function}.
This means that R$^0$ cannot be run with Python 3.

Since Python 2 will soon be discontinued, it may be best to target directly Python 3 and to use the print function (and as a matter of fact, it will work with Python 2).
Another potential problem is the numpy `choice` function that has been introduced only in numpy v1.7.
If for some reason, the installed version of numpy is older, the program will also fail.
One quick fix is to use a simpler way to generate the steps (even if it is less efficient).
Last and most important, we need to specify all the dependencies of our program.

@@

\begin{code}{Random walk (R$^2$)}
# Random walk (R2: repeatable)
# Tested with Python 3
import random
from itertools import accumulate

# Random number generator initialization
random.seed(1)

# Compute random walk for 10 steps
steps = random.choices([-1,+1], k=10)
x = list(accumulate(steps))
print(x) 
\end{code}

\clearpage
\section*{Reproducible (R$^{\mathbf 3}$)}
As nicely explained by Mesnard and Barba \cite{Mesnard:2016}, reproducibility is harder than you may think.

@@

\begin{code}{Random walk (R$^3$)}
# Random walk (R3: reproducible)
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
# Release under the BSD 2-clause license
# Tested with Python 3.6 / macOS 10.12.4 / 64 bits architecture
import random
from itertools import accumulate

@@

if __name__ == '__main__':
    # Unit test
    random.seed(1)
    assert walk(10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]

    # Actual results
    random.seed(2)
    print(walk(10))
\end{code}

@@

\begin{code}{Random walk (R$^4$)}
# Random walk (R4: replicable)
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
# Release under the BSD 2-clause license
# Tested with Python 3.6 / Numpy 1.12.0 / macOS 10.12.4 / 64 bits architecture
import random
import numpy as np

def walk(rng, n):
    """ Random walk for n steps """

    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
    return steps.cumsum().tolist()

def rng(seed):
    """ Return a random number generator initialized with seed """

    rng = random.Random()
    rng.seed(seed)
    _, keys, _ = rng.getstate()
    rng = np.random.RandomState()
    state = rng.get_state()
    rng.set_state((state[0], keys[:-1], state[2], state[3], state[4]))
    return rng

if __name__ == '__main__':
    # Unit test
    assert walk(rng(seed=1), 10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]

    # Actual results
    print(walk(rng(seed=2), 10))
\end{code}

@@

\begin{code}{Random walk (R$^5$)}
# Random walk (R5: re-usable)
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C.Y. Benureau
# Release under the BSD 2-clause license
# Tested with Python 3.6 / Numpy 1.12.0 / macOS 10.12.4 / 64 bits architecture
import random
import numpy as np

def walk(rng, n):
    """ Random walk for n steps """

    steps = 2*(rng.uniform(-1,+1,n) > 0) - 1
    return steps.cumsum().tolist()

def rng(seed):
    """ Return a random number generator initialized with seed """

    rng = random.Random()
    rng.seed(seed)
    _, keys, _ = rng.getstate()
    rng = np.random.RandomState()
    state = rng.get_state()
    rng.set_state((state[0], keys[:-1], state[2], state[3], state[4]))
    return rng

if __name__ == '__main__':
    # Unit test
    # assert walk(rng(seed=1), 10) == [-1, 0, 1, 0, -1, -2, -1, 0, -1, -2]

    import argparse
    parser = argparse.ArgumentParser("Random walk")
    parser.add_argument('--seed', type=int, default=1,
                        help='Seed for random number generator ')
    parser.add_argument('n', type=int, default=10,
                        help='Number of step(s) to walk')
    args = parser.parse_args()

    print("Seed:", args.seed)
    print("Number of steps:", args.n)
    print("Result:",  walk(rng(args.seed), args.n))
\end{code}
